---
title: "2020 Debate Transcripts"
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float: yes
author: "**Daniel Zaltz**"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

```{r setup, include=FALSE, warnings=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
The debate transcripts are made available online through a variety of platforms. I have
not searched for pre-processed data and instead make use of the `rvest` package to scrape
data from https://www.rev.com/. For cleaning, I use a variety of functions from `dplyr` 
and `lubridate`. I am working backwards and so present a full description of how I gather
and clean the data using the Vice Presidential debate. I repeat this with the first 
Presidential debate but with fewer comments.

# Data 
## Vice Presidential Debate 10/7/2020 
### Scrape data from web

```{r message=FALSE}
library(tidyverse) # will use a variety of wrangling/cleaning functions
library(rvest) # web scraping package

# read website data
url <- read_html("https://www.rev.com/blog/transcripts/kamala-harris-mike-pence-2020-vice-presidential-debate-transcript")

# extract relevant text from website
transcript_VP1 <- url %>% 
  html_nodes("#transcription p") %>% 
  html_text() %>% 
  tibble()
```
*Note: I used the function `html_nodes()` to focus on the portion of the website that 
contains the text I want to extract. I use a point-and-click CSS selector as a Chrome 
extension, downloaded from https://selectorgadget.com/. The transcript is contained within
the tag "#transcription p".

### Examine raw data
The data is gathered into one column that includes the speaker, time stamp, and transcript.
```{r}
head(transcript_VP1)
```

### Split into relevant columns
First, I separate the speaker/time stamp from the actual transcript. This is fairly simple
because the speaker and time stamp are separated from the text by a new line.
```{r}
transcript_VP1 <- transcript_VP1 %>% 
  separate(
    1, # select first column (in this case, the only column)
    into = c("A", "transcript"), # separate into two columns (temporary names)
    sep = "\n" # separated by new lines, denoted via regex \n
  ) 

head(transcript_VP1)
```
  
The transcript is now in its own column, but I have to split the speaker and time stamp.
This will also be fairly simple because the speaker name is immediately followed by a
colon before showing the time stamp.

```{r}
transcript_VP1 <- transcript_VP1 %>%  
  separate(
    1, # select first column (which contains both speaker and time)
    into = c("speaker", "time"), # separate into new columns
    sep = ": " # separate by the colon after the speaker's name
  )

head(transcript_VP1)
```
  
We now have three columns, one for speaker, one for time, and one for the transcript.

### Clean the data
We should have three speakers (moderator, VP Pence, Senator Harris) and a continuous time,
as there were no breaks in the debate.
  
First, check to see if we have three speakers. From the table we see that there are two
different labels for Kamala Harris. There may be some extra spaces around her name, so we
remove all extra space and re-evaluate. This works, and we now have three speakers.
```{r}
table(transcript_VP1$speaker)
transcript_VP1$speaker <- str_squish(transcript_VP1$speaker)
table(transcript_VP1$speaker)
```
  
The speaker names are long. We can shorten them for easier manipulation later on.
```{r}
transcript_VP1 <- transcript_VP1 %>% 
  mutate(
    speaker = case_when(speaker == "Kamala Harris" ~ "KH",
                        speaker == "Mike Pence" ~ "MP",
                        speaker == "Susan Page" ~ "moderator")
  )
```
  
Now on to the time stamps. These are currently stored as character strings. We may be
interested in analyzing changes throughout the debate so we should create a more useful
class of time stamps. Here is where I make use of the functions `lubridate::hms()` and its
analogue `lubridate::ms()`. The functions convert character strings to hours:minutes:sec
or just minutes:sec. I am not very familiar with these functions and ran into some issues
at first -- the `hms()` command gave me trouble when it was evaluating a time that only
had minutes and seconds. To get around this, I implement a simple logical expression. If
the time stamp has one colon it means there is just minutes:sec and therefore I will use
the `ms()` function, but if it has two colons there must be hours:minutes:sec and I then 
use the `hms()` function. I would love some feedback on a way around this if there is one.

```{r warning=FALSE, message=FALSE}
transcript_VP1 <- transcript_VP1 %>% 
  mutate(
    time = case_when(
      str_count(time, pattern = ":") == 1 ~ lubridate::ms(time),
      str_count(time, pattern = ":") == 2 ~ lubridate::hms(time))
  )
```
  
Now I check to see how we did when converting from character to time. The debate was one
uninterrupted event and our time stamps should reflect this -- I test this with a simple
plot.
```{r fig.align="center", fig.height=5, fig.width=5}
plot(transcript_VP1$time)
```
  
Alas! Something has gone wrong. This should be one straight line. The x-axis shows the 
index (or row number) of where the time seems to restart from the beginning. From the
graph this seems to occur somewhere around row 130. Let's inspect.
```{r}
transcript_VP1[130:140,]
```
  
As you can see, there is a break in time between rows 135 and 136. Why? I go back to the
website and see that they divided their transcription into two parts, but the actual
debate did not have any intermissions. We need to edit the time stamps. This takes a
little bit of sleuthing:

1. The last time stamp from part 1 of the transcription is 38:16.
2. Re-watch the video that lines up with this part of the transcript. The moderator speaks for appx. 9s.
3. Pence responds immediately. This is the first item from part 2 of the transcript. The time stamp for his response is 02:09, but in reality it should be 00:09 (corresponding to 9s into part 2). Therefore, all time stamps in part 2 are off by 02:00.
4. We need to subtract 02:00 from all time stamps in part 2 and then re-add back the time that had already accrued in part 1 (38:16).
5. Therefore, to fully correct all time stamps from part two, we add an aggregate of 36:16.

```{r}
transcript_VP1$time[136:nrow(transcript_VP1)] <- 
  transcript_VP1$time[136:nrow(transcript_VP1)] + lubridate::ms("36:16")
```
  
  
Check to see if the time appears to follow a continuous sequence.
```{r fig.align="center", fig.height=5, fig.width=5}
plot(transcript_VP1$time)
```
  
Now, we inspect the time stamps, focusing on those surrounding this arbitrary part 1/2 divide.
```{r}
transcript_VP1$time[134:138]
```
  
Some of the seconds exceed 60. Oops. Let's change this by using `roll = TRUE`.
```{r warning=FALSE, message=FALSE}
transcript_VP1$time <- lubridate::ms(transcript_VP1$time, roll = TRUE)
```
  
Upon further inspection things look good but I get an NA for the first time stamp, which
should be 00:00. Add that manually.
```{r}
transcript_VP1$time[1] <- lubridate::ms("00:00")
head(transcript_VP1)
```

**UPDATE:** The rogue apostrophe disaster. Yikes. This was a tough one. After initially
cleaning the data and starting to run some descriptive analyses I was having a hard time
replacing all contractions (I'm, you're, etc.) with full words (I am, you are, etc). I
couldn't figure out why until, finally, I realized that the apostrophes from the transcript
are different than the apostrophes that are recognized by `textclean::replace_contraction()`.


```{r}
transcript_VP1 <- transcript_VP1 %>% 
  mutate(transcript = str_replace_all(transcript, "’", "'"))
```
  


## Presidential Debate 9/29/2020
This follows a similar process as above and as such I have not included as much commentary.
### Scrape data from web
```{r}
url <- read_html("https://www.rev.com/blog/transcripts/donald-trump-joe-biden-1st-presidential-debate-transcript-2020")

transcript_P1 <- url %>% 
  html_nodes("#transcription p") %>% 
  html_text() %>% 
  tibble()
```
### Examine raw data
Again we have arbitrary parts 1 and 2. In this transcript, there is a blank line at row 180.
```{r}
transcript_P1 <- transcript_P1[-180,] # remove blank line denoting the beginning of part 2
```
  
### Split into relevant columns
```{r}
transcript_P1 <- transcript_P1 %>% 
  separate(
    1,
    into = c("A", "transcript"),
    sep = "\n"
  ) 
```

There is also no timestamp at the first row of part 2. Add that here.
```{r}
transcript_P1$A[180] <- "Chris Wallace: (00:00)"
```

Continue to split columns.
```{r}
transcript_P1 <- transcript_P1 %>%  
  separate(
    1,
    into = c("speaker", "time"),
    sep = ": "
  )
```

### Clean the data
Rename variables and change time stamp class.
```{r message=FALSE, warning=FALSE}
transcript_P1 <- transcript_P1 %>% 
  mutate(
    speaker = as.factor(case_when(speaker == "Chris Wallace" ~ "moderator",
                        speaker == "President Donald J. Trump" ~ "DT",
                        speaker == "Vice President Joe Biden" ~ "JB")),
    time = case_when(str_count(time, pattern = ":") == 1 ~ lubridate::ms(time),
                     str_count(time, pattern = ":") == 2 ~ lubridate::hms(time))
  )

```

  
Adjust part 2 time following similar method shown in prior section.
```{r}
transcript_P1$time[180:nrow(transcript_P1)] <- 
  transcript_P1$time[180:nrow(transcript_P1)] + lubridate::ms("24:30")
head(transcript_P1)
```
  

**UPDATE:** Apostrophe-aclypse 2020 averted. See above for explanation.
```{r}
transcript_P1 <- transcript_P1 %>% 
  mutate(transcript = str_replace_all(transcript, "’", "'"))
```
  
